---
title: "SOK-1005 Arbeidskrav 2"
author: "12"
format: pdf
echo: true
output: true
editor: visual
warning: false
---

# Assignment - 2: Scraping an html table from a web page (Compulsory Assignment).

```{r, message=FALSE, warning=FALSE}
rm(list=ls()) 
library(tidyverse)
library(janitor)
library(lubridate)
library(zoo)
library("rvest")
```

## Task 1

```{r, message = FALSE, warning=FALSE}
content <- read_html("https://www.motor.no/aktuelt/motors-store-vintertest-av-rekkevidde-pa-elbiler/217132")
tables <- content %>% html_table(fill = TRUE)
df_car <- tables[[1]]
df_car <- df_car[-1,]
```

```{r, message = FALSE, warning=FALSE}
df_car$X3 <- gsub("km", "", as.character(df_car$X3))
df_car <- df_car %>%
  rename(model=X1) %>% 
  rename(stop=X3) %>% 
  rename(wltp=X2) %>% 
  rename(deviation=X4) %>% 
  mutate(stop=as.numeric(stop))
df_car[is.na(df_car)] <- 0

df_car$wltp <- gsub("km.*", "", as.character(df_car$wltp))

df_car <- df_car %>% 
  filter(stop != 0) %>% 
  mutate(wltp=as.numeric(wltp))

```

```{r, message=FALSE, warning=FALSE}
df_car %>% 
  ggplot(aes(x=wltp, y=stop))+
  geom_point()+
  geom_abline(coef=c(0,1), col="red", size=1, lwd=1)+
  geom_text(x=350,y=420, label="Expected range", color="red")+
  labs(x="Stated range (km)", y="Actual mileage (km)",
       title="Range test results of electric cars", subtitle = "31 electric car models tested by Motor")+
  xlim(200,600)+
  ylim(200,600)+
  theme_bw()
```

## Task 2

```{r}
lm(stop ~ wltp, data = df_car)
```

The output provides the estimated coefficients for the intercept and the **`wltp`** variable. The intercept is estimated to be -26.6450, which represents the estimated value of **`stop`** when **`wltp`** is zero. The slope coefficient for **`wltp`** is estimated to be 0.8671, which means that for every one unit increase in **`wltp`**, the estimated value of **`stop`** increases by 0.8671.

Overall, this output tells us about the linear relationship between **`stop`** and **`wltp`**, and provides estimates for the intercept and slope coefficients that can be used to make predictions about **`stop`** based on values of **`wltp`**.

```{r, message=FALSE, warning=FALSE}
df_car %>% 
  ggplot(aes(x=wltp, y=stop))+
  geom_point()+
  geom_abline(coef=c(0,1), col="red", size=1, lwd=1)+
  geom_smooth(method=lm, se=FALSE)+
  geom_text(x=350,y=420, label="Expected range", color="red")+
  geom_text(x=400,y=280, label="Avarage", color="blue")+
  labs(x="Stated range (km)", y="Actual mileage (km)",
       title="Range test results of electric cars", subtitle = "31 electric car models tested by Motor")+
  xlim(200,600)+
  ylim(200,600)+
  theme_bw()
```
